---
title: "Practice #1: Exploring R and RStudio"
date: Spatial Data Analysis and Visualization with R
output: 
  html_document: 
    css: css/plain_html.css
    includes:
      in_header: header.html
    lib_dir: lib
    self_contained: no
    smart: no
    theme: lumen
    toc: yes
    toc_float: yes
---

## Open R Studio

<div class="indented1">
- explore the different windows
- how to change font size and colors
- running commands from the script window
- packages tab
- help menu, package help  
</div>

## Use R as a calculator

```{r echo=TRUE, eval=FALSE}
2 + 3
56 / 5
2 ^ 3
(4 + 5) * 7
sin(pi / 2)
```

<div class="indented1">
&#9758; When it comes to math expressions, R works a lot like Excel  
&#9758; R generally ignores white space  
</div>

## Save results to variables

```{r echo=TRUE, eval=FALSE}
x = 3 + 4
x
shoe_size = 9.5
shoe_size
Shoe_Size
```

<div class="indented1">
&#9758; Variable names can be long  
&#9758; Avoid using special characters in a variable name  

<div style="margin-left:2em;">
- numbers ok, but not at the beginning. `data1` works. `1data` doesn't.
- use underscores instead of dashes
</div>

&#9758; Variable names are case sensitive
</div>

## Use functions

```{r eval=FALSE}
sqrt(125)
round(x=3.14159, digits=2)
runif()
runif(n=10)
y= seq(from=20, to=40, by=2)
y
palette()
```

<div class="indented1">
&#9758; **All** functions require parentheses ()  
&#9758; Some functions require you to put arguments in the parentheses, some don't  
&#9758; Multiple arguments have to be separated by commas  
&#9758; R functions can return more than one element  
&#9758; R variables can store more than one element (aka *vector*)  
&#9758; Use the up & down arrow keys to bring back your history  
</div>

## Getting Help 

```{r eval=FALSE}
?round
help(seq)
help(package="stats")
browseVignettes("raster")
```

<div class="indented1">
&#9758; **All** functions have a help page  
&#9758; Many packages have a help page  
&#9758; Many packages have tutorials or workflow examples in Vignettes   
</div>

## Install Packages

Install the `leaflet` package

<div class="indented1">
&#9758; you can install packages in RStudio using the *Packages* tab
&#9758; you have to install (download) packages before you can use them  
</div>

## Use the Script Window

Copy-paste the following into a new script.

```{r eval=F}
# Happy Saint Patricks Day!
plot(NULL, xlim=c(-1,1), ylim=c(-2,1), axes=FALSE, asp=1, xlab="", ylab="")
x <- seq(from=-2, to=0, by=0.1)
y <- -(1/8) * x^2 - 0.01
points(-1*y,x,type="l", col="darkgreen", lwd=4)
theta <- seq(from=0, to=2*pi, by=0.05)
r <- sin(2*theta) + 0.25 * sin(6*theta)
points(x = r * cos(theta), y = r * sin(theta), type="l", col="darkgreen", lwd=3)
```

<div class="indented1">
&#9758; You can write code in a script to save and reuse it  
&#9758; Execute code line-by-line by clicking the Run button or ctrl+enter  
&#9758; Any line that starts with '#' is interpreted as a comment and ignored  
</div>

Let's make an interactive map using the `leaflet` package. Copy the following into a script and run it.

```{r eval=FALSE}
# Let's make an interactive map of a sample dataset of earthquakes 
library(leaflet)
m = leaflet(quakes) %>% addTiles() %>% addMarkers(~long, ~lat)
m
```

<div class="indented1">
&#9758; R comes with some sample datasets (like `quakes`)  
&#9758; Before you can use an installed package, you have to load it into memory using `library()`
&#9758; Learn more about leaflet from [RStudio](https://rstudio.github.io/leaflet/){target="_blank"} or this [UC Berkeley workshop](https://github.com/dlab-geo/r-leaflet-workshop){target="_blank"}.
</div>


## Plot some data

```{r}
y = rnorm(500)
hist(y)
hist(y, breaks=20, col="grey", main="Bell Curve")
x = seq(from=0, to=4*pi, by=0.1)
plot(x, cos(x))

library(sp)
library(rnaturalearth)
plot(countries110)
```

<div class="indented1">
&#9758; R has a number of basic, built-in plotting functions  
&#9758; most plotting functions have optional arguments to cumtomize the look  
&#9758; when you pass text to a function, you have to enclose it in quotes  
&#9758; there are packages which provide more plotting options   
</div>

