---
title: Working with 2D and Spatial Data
output: 
  slidy_presentation:
    css: css/slidy_styles.css
    df_print: default
    duration: 0
    font_adjustment: 2
    footer: "Data Analysis and Visualization with R | <a href='https://ajlyons.github.io/sfdata/'>https://ajlyons.github.io/sfdata/</a>"
    self_contained: no
    lib_dir: lib
    smart: no
    includes:
      in_header: header.html
    increment: no
---

# Two-dimensional data

<script language="javascript" type="text/javascript">w3c_slidy.mouse_click_enabled = false;</script>

R has two object types for handling tabular data: **matrices** and **data frames**.

```{r}
head(airquality)
```

There are number of sample datasets that come with R, many of which are data frames. Functions to explore data frames include <tt>names()</tt>, <tt>head()</tt>, <tt>nrow()</tt>, <tt>ncol()</tt>, <tt>dim()</tt>.

```{r}
#View(mtcars)
names(mtcars)
head(mtcars)
nrow(mtcars)
ncol(mtcars)
```

You can grab an individual column by name with the <tt>$</tt> selector. 

```{r plot_wt_mpg, fig.align='center', cache=TRUE}
summary(mtcars$wt)
plot(x=mtcars$wt, y=mtcars$mpg, pch=16, main="Gas Mileage vs Weight")
```

# Filtering Rows and Columns

You can reference specific rows specific rows and columns with square bracket notation. Only this time we need two expressions separated by a comma.

<tt>mydataframe[</tt> *rows* , *cols* <tt>]</tt>

The expression for *rows* & *cols* can be integers (row and column indices), or Booleans. In addition, with data frames you can use a vector of column names to specific the columns. To return all the rows or all the columns, omit the expression. 

```{r}
library(maps)
data(canada.cities)
head(canada.cities)
```

Columns can be specified either by name or index number. The following two expressions are equivalent (note how the expression for *rows* has been omitted, so all rows are returned). 

```{r}
head(canada.cities[,1])
head(canada.cities[,"name"])
```

The *rows* expression can also be an expression that returns logicals.

```{r plot_big_cities, small.mar=TRUE, fig.align='center', cache=TRUE}
idx <- (canada.cities$pop > 100000)
table(idx)
big_cities <- canada.cities[idx,   ]
plot(big_cities[ , c("long","lat")], asp=1, pch=16  )
```

# Working with Files and Folders

To import and export data, you have to 


# Importing a csv

Whenever you import or export data, 

Setting up 

RStudio has a 'import dataset' tool in the upper right pane. 

```{r import_evictions}

```



